Complete representability for canonical extensions of representable cylindric algebras

### 2021-11-21 cylindric-notes/Text.tex
more updates on RCA_omega

Given $\mathcal{C} \in {\bf RCA}_{\omega}$, whether $\mathcal{C}^{+}$ has a complete, $\omega$-dimensional representation? \cite{hirsch2002relation}

\section{Atomic Representations}

@@

\begin{defin} Let $\mathcal{F} = \langle W, (R_{i})_{i \in I} \rangle$ is an inner substructure \footnote{Or alternatively, a generated subframe} of $\mathcal{F}' = \langle W', ({R'}_{i})_{i \in I} \rangle$, if $W \subseteq W'$ and the embedding $\mathcal{F} \hookrightarrow \mathcal{F}'$ is a bounded morphism.
\end{defin}

@@

Let $\alpha$ be an ordinal. Let $\prescript{\alpha}{}U$ be the set of all functions mapping $\alpha$ to a non-empty set $U$. We denote $x(i) = x_i$ for
$x \in \prescript{\alpha}{}U$ and $i < \alpha$.

@@

Let $\mathcal{A} \in {\bf CA}_{\omega}$, then $\mathcal{A}$ has a \emph{complete representation}, if its representation preserves all existing suprema. In other words, $\mathcal{A}$ is completely representable.

@@

The following definition of an $\omega$-frame is due to \cite{Venema2013}.
\begin{defin}
  A cylindric $\omega$-frame is a structure $\mathcal{F} = \langle W, (R_i)_{i < \omega}, (E_{ij})_{i, j < \omega} \rangle$ where $(R_i)_{i < \omega}$ are binary relations and $(E_ij)_{i, j < \omega}$ are unary relations such that, for all $i, j, k < \omega$:
  \begin{enumerate}
  \item Every $R_i$ is an equivalence relation on $W$,
  \item $R_i \circ R_j = R_j \circ R_i$, that is, the set $(R_i)_{i < \omega}$ forms a commutative semigroup under composition.
  \item For all $x \in W$, $E_{ii}(x)$ holds.
  \item For all $x, y, z \in W$, $x R_i y \: \& \: E_{ij}(y) \: \& \: x R_i z \: \& \: E_{ij}(y)$ implies $y = z$.
  \item For all $x \in W$, $E_{ij}(x)$ iff there exists $y \in W$ such that $x R_k y$, $E_{ik}(y)$, and $E_{kj}(y)$.
  \end{enumerate}
\end{defin}

@@

\begin{defin}
Let $U$ be a set and $V \subseteq \prescript{\omega}{}U$ be a non-empty subset of the full Cartesian space of dimension $\omega$, then an $\alpha$-dimension Cartesian structure generated by $V$ is an $\omega$-frame $\mathfrak{G}(V) = \langle V, (R_{i})_{i < \omega}, (E_{ij})_{i, j < \omega} \rangle$ such that:
\begin{enumerate}
\item $R_i = \{ (w, v) \: | \: w, v \in V, w_k = w_k, k < \omega, i \neq k \}$
\item $E_{ij} = \{ w \in V \: | \: w_i = w_j \}$
\end{enumerate}
$\mathfrak{G}(\prescript{\omega}{}U)$ is the full $\omega$-dimensional Cartesian structure. $\mathcal{F} \mathfrak{c}\mathfrak{t}_{\omega}$ is the class of all full $\omega$-dimensional Cartesian structures.
\end{defin}

Every cylindric set algebra is a subalgebra of some complex algebra induced by an $\omega$-dimensional Cartesian structure. That is, we have the following characterisation:

@@

In this section, we reproduce the results related to characterisation ${\bf RCA}_{\omega}$. The following results are due to Goldblatt \cite{goldblatt1995elementary}.

\begin{lemma}
${\bf RCA}_{\omega} = {\bf S} \: \mathfrak{Cm} {\bf Ud} \mathcal{F} \mathfrak{ct}_{\omega}$
\end{lemma}

@@

>>>>

\begin{itemize}
\item Given $\mathcal{C} \in {\bf RCA}_{\omega}$, whether $\mathcal{C}^{+}$ has a complete, $\omega$-dimensional representation? The conjecture is yes. \cite{hirsch2002relation}
\item Whether ${\bf RCA}_{\omega}$ is barely canonical. The conjecture is yes.
\end{itemize}
\section{Atomic Representations}

@@

\begin{defin} Let $\mathcal{F} = \langle W, (R_{i})_{i \in I} \rangle$ is an inner substructure \footnote{Or alternatively, a generated subframe} of $\mathcal{F}' = \langle W', ({R'}_{i})_{i \in I} \rangle$, if $W \subseteq W'$ and the embedding $\mathcal{F} \hookrightarrow \mathcal{F}'$ is a bounded morphism. Let $\mathbb{F}$ be a class atom structures, then $\mathbb{S}(\mathbb{F})$ is the closure of $\mathbb{F}$ under generated subframes.
\end{defin}

@@

Let $\alpha$ be an ordinal. Denote $\{ f \: | \: f \alpha \to U\}$ as $\prescript{\alpha}{}U$. $x_i$ stands for $x(i)$, where
$x \in \prescript{\alpha}{}U$ and $i < \alpha$.

@@

Let $\mathcal{A} \in {\bf CA}_{\omega}$, then $\mathcal{A}$ has a \emph{complete representation}, if its representation preserves all existing suprema. In other words, $\mathcal{A}$ is \emph{completely representable}.

@@

The following definition of an $\omega$-frame is due to \cite{Venema2013}.
\begin{defin}
  A cylindric $\omega$-frame is a structure $\mathcal{F} = \langle W, (R_i)_{i < \omega}, (E_{ij})_{i, j < \omega} \rangle$ where $(R_i)_{i < \omega}$ are binary relations and $(E_ij)_{i, j < \omega}$ are unary relations such that, for all $i, j, k < \omega$:
  \begin{enumerate}
  \item Every $R_i$ is an equivalence relation on $W$,
  \item $R_i \circ R_j = R_j \circ R_i$, that is, the set $(R_i)_{i < \omega}$ forms a commutative semigroup under composition.
  \item For all $x \in W$, $E_{ii}(x)$ holds.
  \item For all $x, y, z \in W$, $x R_i y \: \& \: E_{ij}(y) \: \& \: x R_i z \: \& \: E_{ij}(y)$ implies $y = z$.
  \item For all $x \in W$, $E_{ij}(x)$ iff there exists $y \in W$ such that $x R_k y$, $E_{ik}(y)$, and $E_{kj}(y)$.
  \end{enumerate}
  $\mathcal{C}\mathfrak{a}_{\omega}$ is the class of all $\omega$-frames.

  If $\mathcal{F} \in \mathcal{C}\mathfrak{a}_{\omega}$ and $x \in \mathcal{F}$, then $\mathcal{F}^{x}$ is a generated subframe generated by $x$, which is defined standardly. Generally, $\mathcal{F}_1$ is a generated subframe of $\mathcal{F}_2$, if $\underline{\mathcal{F}_1} \subseteq \underline{\mathcal{F}_2}$ and $\underline{\mathcal{F}_1}$ is closed under under ${R_i}_2$ equivalences for every $i < \omega$. That is:
  \begin{center}
  For all $i < \omega$ and $x \in \mathcal{F}_1$, we have ${R_i}_2(x) \subseteq \mathcal{F}_1$ and, thus, ${R_i}_1(x) = {R_i}_2(x)$.
  \end{center}
\end{defin}

We have the following connection betweens $\omega$-frames and their generated subframes, which is standard for modal logic:
\begin{prop}
Let $\mathcal{F} \in \mathcal{C}\mathfrak{a}_{\omega}$, then
\begin{enumerate}
\item $\mathcal{F} = \coprod \limits_{x \in \mathcal{F}} \mathcal{F}^{x}$,
\item $\mathfrak{Cm}(\mathcal{F}) \cong \prod \limits_{x \in \mathcal{F}} \mathfrak{Cm}(\mathcal{F}^{x})$,
\item $\mathfrak{Cm}(\mathcal{F}^{x})$ is subdirectly irreducible.
\end{enumerate}
\end{prop}

It is known that $\mathcal{C}\mathfrak{a}_{\omega}$ forms an elementary class, since one can express the conditions of an $\omega$-frame with the first-order language.

@@

\begin{defin}
Let $U$ be a set and $V \subseteq \prescript{\omega}{}U$ be a non-empty subset of the full Cartesian space of dimension $\omega$, then an $\alpha$-dimension Cartesian structure generated by $V$ is an $\omega$-frame $\mathfrak{S}(V) = \langle V, (R_{i})_{i < \omega}, (E_{ij})_{i, j < \omega} \rangle$ such that:
\begin{enumerate}
\item $R_i = \{ (w, v) \: | \: w, v \in V, w_k = w_k, k < \omega, i \neq k \}$
\item $E_{ij} = \{ w \in V \: | \: w_i = w_j \}$
\end{enumerate}
$\mathfrak{S}(\prescript{\omega}{}U)$ is the full $\omega$-dimensional Cartesian structure. $\mathcal{F} \mathfrak{c}\mathfrak{t}_{\omega}$ is the class of all full $\omega$-dimensional Cartesian structures.
\end{defin}

Clearly $\mathcal{F} \mathfrak{c}\mathfrak{t}_{\omega} \subseteq \mathcal{C}\mathfrak{a}_{\omega}$.

We have the following connection between ${\bf RCA}_{\omega}$, ${\bf IGs}_{\omega}$, and complex algebras of full Cartesian structures:
\begin{center}
${\bf RCA}_{\omega} = {\bf IGs}_{\omega} = {\bf S} \mathfrak{Cm} {\bf Ud} \mathcal{F} \mathfrak{c}\mathfrak{t}_{\omega} = {\bf S P} \mathfrak{Cm}  \mathcal{F} \mathfrak{c}\mathfrak{t}_{\omega}$.
\end{center}
This follows from the fact that ${\bf Cs}_{\omega} = \mathfrak{Cm} \mathcal{F} \mathfrak{c}\mathfrak{t}_{\omega}$. Every generalised cylindric set algebra is a subdirect product of cylindric set algebras, thus, a generalised cylindric set algebra is a complex algebra of disjoint union of some full Cartesian spaces. But ${\bf RCA}_{\omega}$ is the closure of ${\bf Cs}_{\omega}$ under isomorphism.

\begin{defin}
The weak Cartesian space with base $U$ and dimension $\omega$ determined by $x \in \prescript{\omega}{}U$ is the set:
\begin{center}
$\prescript{\omega}{}U^{(x)} = \{ y \in \prescript{\omega}{}U \: | \: |\{ k < \omega \: | \: x_k \neq y_k \}| < \aleph_0 \}$
\end{center}
$\mathfrak{S}(\prescript{\omega}{}U^{(x)})$ is a weak Cartesian structure of dimension $\omega$. $\mathcal{W} \mathfrak{ct}_{\omega}$ is the class of all weak Cartesian structure of dimension $\omega$ up to isomorphism.
\end{defin}
Note that we have $\mathcal{W} \mathfrak{ct}_{\omega} \subseteq \mathcal{C}\mathfrak{a}_{\omega}$.

Every cylindric set algebra is a subalgebra of some complex algebra induced by an $\omega$-dimensional Cartesian structure. In other words,

@@

In this section, we reproduce the results related to characterisation ${\bf RCA}_{\omega}$. The following results are due to Goldblatt \cite{goldblatt1995elementary}. This denotes that a cylindric algebra of dimension algebra is representable iff it is isomorphic to a subalgebra of the complex algebra of disjoint sum of some full $\omega$-dimensional Cartesian structure. Assuming the duality, this is equivalent to the standard definition of representability formulated in terms of sublagebras of subdirect products.

\begin{lemma} \label{rcachar}
${\bf RCA}_{\omega} = {\bf S} \: \mathfrak{Cm} \mathbb{S} {\bf Ud} \mathcal{F} \mathfrak{ct}_{\omega} = {\bf S} \: \mathfrak{Cm} \mathbb{S} {\bf Ud} \mathcal{W} \mathfrak{ct}_{\omega} = {\bf IGws}_{\omega}$
\end{lemma}

\begin{proof}

\end{proof}

Here we use the following fact related to canonical varieties generated by some class of complex algebras. Let ${\bf K}$ be an elementary class of relational structures, then:
\begin{center}
If ${\bf K}$ is closed under $p$-morphic images, generated subframes, and disjoint unious, \\ then ${\bf S} \mathfrak{Cm} {\bf K}$ is a canonical variety.
\end{center}
One may think of this fact a more abstract version of Fine's theorem which claims that every elementary modal logic is canonical \cite{fine1975some}. This version denotes the same fact, but it is formulated in terms of varieties BAOs generated by complex algebras of some atom structures.

This is a specialised version of \cite[Theorem 4.4]{goldblatt1995elementary} formulated for dimension $\omega$.

@@

\begin{proof}
By Lemma~\ref{rcachar},
\end{proof}
